---
title: "In-Class Exercise 1 Building Elegant Graphics with GGPLOT2"
description: "In-Class Exercise 1"
author:
  - name: Clement Ong 
    url: https://example.com/norajones
    affiliation: ISSS608 Visual Analytics
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Install and Loading the Required Libraries 
```{r echo=TRUE}
packages = c('tidyverse', 'ggrepel')
for (p in packages){
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

Why Character.only :  If you look at help to library (?library) then you got (in Argument section): "package the name of a package, given as a name or literal character string, or a character string, depending on whether character.only is FALSE (default) or TRUE." 

Note that for the above warnings, ***dplyr::filter()*** for a specific kind of filter. To be more careful, can ***dplyr::*** and wait for the suggestions 

To understand more about the tidyverse : 
- https://www.tidyverse.org 
- Example of Reference Docs : https://readr.tidyverse.org 

# Import Data into R

```{r}
exam_data <- read_csv("data/Exam_data.csv") # you can do a filter or sort, in df environement
```

Differences between Read.csv and Read_csv
- First.Name vs First Name 

---

# Plotting First Chart with GGPLOT2


##### Plotting a grpah in GGPLOT has an order:
- Data : DF (blank face means data loaded)
- Aes : X or Y or BOTH
- Geom : geon_point(), geom_line(), geom_boxplot(), 
- Facets : Creating the column of the chart (PANE)
- Statistics : median, average ... 
- Coordinates : cartesian (x and y axis)
- Theme : Change themes for better colors 

```{r echo=TRUE}
ggplot(exam_data, aes(x=MATHS)) +
  geom_histogram(bins=10,boundary=100, color="black",fill="grey") + 
  geom_hline(aes(yintercept = mean(MATHS))) +
  ggtitle("Distribution of Math Scores")
```

Note that for histogram the default is 30 bins. ***Make Sure adjust the bins and the Colors***

Notes : 
- **Boundary is used to ensure the chart does not over run 100**
- you can add more like FILL in the ggplot aesthetics which is used for category (Remove the fill in geom object)
- color = 'grey30' 30 is the intensity of the grey- border of the histogram
- boxplot - notch=TRUE
- boxplot + point
- geom_violin() - as the distribution + geom_boxplot(alpha=0.5) - act as the guiding line 
- Box plot with summary point, stat_summary(geom='point', fun.y='mean') vs geom_point(stat='summary', fun.y='mean')
- leanr to work with facets and coordinates

```{r echo=TRUE}
ggplot(exam_data, aes(x=MATHS, fill=GENDER)) +
  geom_histogram(bins=20,boundary=100, color="grey38") + 
  facet_wrap(~GENDER) +
  ggtitle("Distribution of Math Scores")
  
```








